Pipit: Reactive Systems in \fstar ~(Extended Abstract)

### 2023-05-09 2023pipit/Main.tex
ok

@@

We are interested in verifying an antilock braking system for a motorcycle\cite{huang2010design}.
Antilock brakes are designed to stop wheels from locking up in emergency braking events.
Antilock braking controllers generally work by computing the \emph{wheel slip} for the wheels, which is a ratio of the wheel's rotational speed compared to the vehicle's speed.
To compute the wheel slip, then, the controller must first estimate the vehicle's speed.

@@

\begin{tabbing}
  @MM@\= @let @\= \kill
  @let@ veh\_speed\_estimate $\omega_F$ $\omega_R$ $a_z$ = \\
    \> @let@ $v_F$ = $\omega_F \cdot $ circumference @in@ \\
    \> @let@ $v_R$ = $\omega_R \cdot $ circumference @in@ \\
    \\
    \> @let rec@ $\floor{\hat{v}}$ = @if@ $v_F \approx_\epsilon v_R$ \\
    \> \> @then@ $\min v_F v_R$ \\
    \> \> @else@ $(\xthen{\min v_F v_R}{\xfby{0}{\floor{\hat{v}}}}) + a_z - \epsilon$ @in@ \\
    \> @let rec@ $\ceil{\hat{v}}$ = @if@ $v_F \approx_\epsilon v_R$ \\
    \> \> @then@ $\max v_F v_R$ \\
    \> \> @else@ $(\xthen{\max v_F v_R}{\xfby{0}{\ceil{\hat{v}}}}) + a_z + \epsilon$ @in@ \\
    \\
    \> $(\floor{\hat{v}}, \ceil{\hat{v}})$
\end{tabbing}

Cite: Kind2\cite{champion2016kind2}, Lustre\cite{caspi1995functional}, Vélus\cite{bourke2017formally}.
The original Kind\cite{hagen2008scaling}.

@@

The semantics here, particularly the recursive rule, differ from previous semantics of Lustre.
In the reactive semantics of \cite{caspi1995functional}, as well as the coinductive semantics of Vélus \CITE{bourke2017formally find best paper}, the semantics for mutually recursive definitions is locally non-deterministic, in that the semantics does not define which particular order mutual definitions are evaluated in.
This local non-determinism complicates certain proofs, even though the overall evaluation \emph{is} deterministic \TODO{vague}.

@@

Invariant generation: ICE\cite{garg2014ice}; template-based\cite{kahsai2011instantiation}; PDR explanation\cite{een2011efficient}; IC3 first introduced\cite{bradley2011sat}

@@

We have implemented an unverified imperative code generation transform, by making use of \fstar{}'s existing C code extraction support\cite{protzenko2017verified}.
% This transform generates \fstar{} functions in the

@@

Other compilers for Lustre-style languages, such as Vélus\cite{bourke2017formally} and Heptagon \cite{gerard2012modular} ensure that the generated code can be executed in bounded memory and that each step takes a bounded time.
With a carefully chosen set of primitives, the core streaming operations shown here would ensure bounded memory and time.
For practicality, however, we currently allow embedding arbitrary total functions from the \fstar{} meta-language, which are not necessarily bounded in time.
(Is this the same as Lucid Synchrone and Zélus? \CITE)

>>>>

% \usepackage{cite}

@@

\TODO{Required for submission: writing}
\begin{itemize}
  \item introduction
  \item running example: vehicle speed estimate with ``error tolerance limit'' property ($\blacklozenge_{t} (v_F \approx_{\epsilon} v_R) \implies \lfloor \hat{v}' \rfloor \approx_{t\epsilon} \lceil \hat{v}' \rceil$)?
    + uses modal operators
    + property is k-inductive for small $t$, requires explicit invariant for larger $t$, motivates k-induction
    - not super interesting
    - proof requires CSE, which isn't implemented yet
  \item describe translation to transition system? detail probably won't fit, translation invariant is the most interesting part.
\end{itemize}
\TODO{Required for submission: implementation}
\begin{itemize}
  \item finish proof of system correctness
  \item state end-to-end / capstone theorem that k-inductively verified systems apply to original semantics
  \item implement unverified CSE pass
  \item verify running example, with CSE and k-induction
\end{itemize}

We are interested in verifying an antilock braking system for a motorcycle \cite{huang2010design}.
Antilock brakes are designed to stop wheels from locking up in emergency braking events.
Antilock braking controllers generally work by computing the \emph{wheel slip} for the wheels, which is a ratio of the wheel's rotational speed compared to the vehicle's speed.
To compute the wheel slip, then, the controller must first estimate the vehicle's speed.
When the wheels are travelling freely with no braking pressure and no wheel slip, one can accurately estimate the vehicle's speed by multiplying the wheel rotational speed by the wheel radius.
Unfortunately, once the wheels begin to slip, it becomes harder to accurately estimate the wheel speed.
For this reason, brake controllers often use a combination of sensors.
Some brake controllers \cite{kobayashi1995estimation} incorporate both wheel speed sensors and accelerometer sensors; the readings are combined with a Kalman filter.

\begin{figure}
\begin{tabbing}
  @MM@\= @let @\= \kill
  @let@ veh\_speed\_estimate $\omega_F$ $\omega_R$ $a_z$ = \\
    \> @let@ $v_F$ = $\omega_F \cdot $ radius @in@ \\
    \> @let@ $v_R$ = $\omega_R \cdot $ radius @in@ \\
    \\
    \> @let rec@ $\floor{\hat{v}}$ = @if@ $v_F \approx_\epsilon v_R$ \\
    \> \> @then@ $\min v_F v_R$ \\
    \> \> @else@ $(\xthen{\min v_F v_R}{\xfby{0}{\floor{\hat{v}}}}) + a_z - \epsilon$ @in@ \\
    \> @let rec@ $\ceil{\hat{v}}$ = @if@ $v_F \approx_\epsilon v_R$ \\
    \> \> @then@ $\max v_F v_R$ \\
    \> \> @else@ $(\xthen{\max v_F v_R}{\xfby{0}{\ceil{\hat{v}}}}) + a_z + \epsilon$ @in@ \\
    \\
    \> $(\floor{\hat{v}}, \ceil{\hat{v}})$
\end{tabbing}

\caption{Implementation of \tt{veh\_speed\_estimate}.}\label{f:veh-speed-estimate}
\end{figure}

@@

Cite: Kind2 \cite{champion2016kind2}, Lustre \cite{caspi1995functional}, Vélus \cite{bourke2017formally}.
The original Kind \cite{hagen2008scaling}.

\TODO{\fstar{} $\to \mbox{F}^{\star}$}
\CITEME{2022 thesis on Agda semantics of Lustre?}

@@

The semantics here, particularly the recursive rule, differ from previous semantics of Lustre.
In the reactive semantics of \citet{caspi1995functional}, as well as the coinductive semantics of Vélus \cite{bourke2017formally} (\CITEME{find best paper}), the semantics for mutually recursive definitions is locally non-deterministic, in that the semantics does not define which particular order mutual definitions are evaluated in.
This local non-determinism complicates certain proofs, even though the overall evaluation \emph{is} deterministic \TODO{vague}.

@@

Invariant generation: ICE \cite{garg2014ice}; template-based \cite{kahsai2011instantiation}; PDR explanation \cite{een2011efficient}; IC3 first introduced \cite{bradley2011sat}

@@

We have implemented an unverified imperative code generation transform, by making use of \fstar{}'s existing C code extraction support \cite{protzenko2017verified}.
% This transform generates \fstar{} functions in the

@@

Other compilers for Lustre-style languages, such as Vélus \cite{bourke2017formally} and Heptagon \cite{gerard2012modular} ensure that the generated code can be executed in bounded memory and that each step takes a bounded time.
With a carefully chosen set of primitives, the core streaming operations shown here would ensure bounded memory and time.
For practicality, however, we currently allow embedding arbitrary total functions from the \fstar{} meta-language, which are not necessarily bounded in time.
(Is this the same as Lucid Synchrone and Zélus? \CITE)
